---
title: "Data Collection:<br> Analyzing Productivity"
subtitle: "Aditya Rathod"
author: "Statistics Period 6"
date: "October 1, 2018"
output:
  ioslides_presentation:
    widescreen: true
    transition: faster
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Background

## Introduction

In March 2018:

- Decided to use Emacs' `org-mode`
  - Keeps track of my tasks
  - Keeps track of my working habits
- At the beginning of each task, I "clock in," work and then "clock out" 

My data is a gold mine for me to analyze my productivity

- In this project, I will take this data and perform an observational study on my productivity

## Research Question

On what days of the week am I most productive?

## Data Collection Method
- Data collected using `org-mode`
- An observational study was appropriate: data collected over time
  - Experiment not possible
    - No way to assign which days to "be" productive or "not" productive.
- $n = 83$

## Materials
- Computer with Emacs installed
- Time

## Procedure {.build}

1. Over a period of time, collect data on time spent working using Emacs' `org-mode` program.

2. Aggregate this data each day to produce the `workdata.csv` file:

```{r message=FALSE, warning=FALSE, include=FALSE}
library(readr)
workdata <- read_csv("workdata.csv")
head(workdata)
```

| Date   | DayWeek   | HoursWorked | Hour | FracHour      | Hours       | 
|--------|-----------|-------------|------|---------------|-------------| 
| 30-Mar | Friday    | 4:33        | 4    | 0.55          | 4.55        | 
| 31-Mar | Saturday  | 4:09        | 4    | 0.15          | 4.15        | 
| 1-Apr  | Sunday    | 5:01        | 5    | 0.01666666667 | 5.016666667 | 


## Explanatory and Response Variables
- The **explanatory variable** in this case is the day of the week (the `DayWeek` column above), which is a *categorical variable*.
- The **response variable** is the number of hours worked that day.

# Analysis

## Analysis: Overview
- Time to analyze the data
- This analysis uses R, a professional statistical analysis package
  - While it is overkill for this small project, it's helpful to learn for future use
  
## Analysis: Part 1 {.smaller .build}
- First, we must mean, SD, and median the hours worked per day based on the day of the week:

```{r message=FALSE, warning=FALSE}
library(knitr)
temp <- aggregate(x = workdata, by = list(workdata$DayWeek), FUN = mean, drop = TRUE)
results <- temp[, c("Group.1", "Hours")]
colnames(results)[1] <- "DayWeek"
colnames(results)[2] <- "HoursMean"

temp <- aggregate(x = workdata, by = list(workdata$DayWeek), FUN = sd, drop = TRUE)
results$HoursSD <- temp[, c("Hours")]

temp <- aggregate(x = workdata, by = list(workdata$DayWeek), FUN = median, drop = TRUE)
results$HoursMedian <- temp[, c("Hours")]

knitr::kable(head(results, 7))
```

## Analysis: Explanation of Part 1

- Mondays are a bit of a mixed bag in terms of work, around 4.3 hours
  - The third most productive day!
  - Has a standard deviation of *4.2 hours*!
    - Means I either work **extremely hard** on Mondays or **get no work done on that day** (I usually *do* work).
- Fridays: I barely get any work done
  - The beginning of the weekend
- On the other hand, Sundays are (usually) my most productive days

## Analysis: Part 2
- Let's visualize these results in bar charts:

## Analysis: Part 2 {.flexbox .vcenter .smaller}
```{r echo=TRUE, display=TRUE}
barplot(results$HoursMean, names.arg = results$DayWeek, xlab = 'Day of Week', ylab = 'Hours Worked',
        main = 'Average Productivity by Day', col = rep('lightblue', 7), cex.names = 0.91)
```

## Analysis: Part 2 {.flexbox .vcenter .smaller}
```{r echo=TRUE, display=TRUE}
barplot(results$HoursSD, names.arg = results$DayWeek, xlab = 'Day of Week',
        ylab = 'Deviation from Mean (hours)', main = 'Variation in Productivity by Day',
        col = rep('lightgreen', 7), cex.names = 0.91)
```

## Analysis: Part 2 {.flexbox .vcenter .smaller}
```{r echo=TRUE, display=TRUE}
barplot(results$HoursMedian, names.arg = results$DayWeek, xlab = 'Day of Week',
        ylab = 'Hours Worked', main = 'Median Productivity by Day', 
        col = rep('purple', 7), cex.names = 0.91)
```

# Conclusion
## {}
- I can conclude that my most productive day is Sundays
  - I am least productive on Fridays
- Must account for variation from the daily average productivity
  - Makes my most productive day Thursdays
    - Due to its low $\sigma$
    - Relatively high value when $\mu - \sigma$.


##  {.flexbox .vcenter}
<span style="font-size: 5rem;">Questions?</span>

##  {.flexbox .vcenter}
### Written analysis available at
### https://applecrazy.github.io/statistics